---
title: "JASA"
author: "Jia Ying"
date: "19/12/2019"
output: html_document
---

# Downloading the datasets for this scrip from:
https://github.com/JY-JiaYING/JASA_Paper

# Analysis A: Articulatory dynamics of tongue lateralization
```{r}
# Loading packages
library(gss)
library(ggplot2)

# Loading data: 
dat <- read.csv("C:/Users/Jia/Desktop/JASA/writing/data/dat.csv")
dat$time <- dat$samp*10

for (i in 1:nrow(dat)){
  dat$type[i] <- paste0(dat$vowel[i],".",dat$syllable[i])
}

dat$type <- as.factor(dat$type)

dat$item <- as.factor(dat$item)

for (i in unique(dat$item)){
  test <- dat$delta[dat$item==i]
  if ( max(abs(test))>20){
    dat <- dat[!dat$item==i,]
  }
}

ss.model <- ssanova(deltaz ~ type + time + type:time,data=dat)

grid <- expand.grid(time = seq(0,290,length=60),
                    type = c("ae.coda","ae.onset","ih.coda","ih.onset")
)

grid$fit <- predict(ss.model,newdata=grid,se=T)$fit
grid$se <- predict(ss.model,newdata=grid,se=T)$se.fit


ggplot(grid,aes(x=time,color=type,group=type)) +
  xlab("Time (ms)") + ylab("Normalized tongue lateralization") +
  geom_line(aes(y=fit,color=type),size=1.5) +
  geom_ribbon(aes(ymin=fit-(1.96*se),ymax=fit+(1.96*se),fill=type),alpha=0.5, color="NA")

```

# Analysis B: Acoustic measurements of tongue lateralization vs. F1 frequency
```{r}

dat$time <- dat$samp*10

for (i in 1:nrow(dat)){
  dat$type[i] <- paste0(dat$vowel[i],".",dat$syllable[i])
}

dat$type <- as.factor(dat$type)

dat$item <- as.factor(dat$item)

for (i in unique(dat$item)){
  test <- dat$delta[dat$item==i]
  if ( max(abs(test))>20){
    dat <- dat[!dat$item==i,]
  }
}

ss.model <- ssanova(f1z ~ type + time + type:time,data=dat)

grid <- expand.grid(time = seq(0,290,length=60),
                    type = c("ae.coda","ae.onset","ih.coda","ih.onset")
)

grid$fit <- predict(ss.model,newdata=grid,se=T)$fit
grid$se <- predict(ss.model,newdata=grid,se=T)$se.fit


ggplot(grid,aes(x=time,color=type,group=type)) +
  xlab("Time (ms)") + ylab("Normalized F1 value") +
  geom_line(aes(y=fit,color=type),size=1.5) +
  geom_ribbon(aes(ymin=fit-(1.96*se),ymax=fit+(1.96*se),fill=type),alpha=0.5, color="NA")

```

# Analysis B: Acoustic measurements of tongue lateralization vs. F2 frequency
```{r}
ss.model <- ssanova(f2z ~ type + time + type:time,data=dat)

grid <- expand.grid(time = seq(0,290,length=60),
                    type = c("ae.coda","ae.onset","ih.coda","ih.onset")
)

grid$fit <- predict(ss.model,newdata=grid,se=T)$fit
grid$se <- predict(ss.model,newdata=grid,se=T)$se.fit


ggplot(grid,aes(x=time,color=type,group=type)) +
  xlab("Time (ms)") + ylab("Normalized F2 value") +
  geom_line(aes(y=fit,color=type),size=1.5) +
  geom_ribbon(aes(ymin=fit-(1.96*se),ymax=fit+(1.96*se),fill=type),alpha=0.5, color="NA")

```

# Analysis B: Acoustic measurements of tongue lateralization vs. F3 frequency
```{r}
ss.model <- ssanova(f3z ~ type + time + type:time,data=dat)

grid <- expand.grid(time = seq(0,290,length=60),
                    type = c("ae.coda","ae.onset","ih.coda","ih.onset")
)

grid$fit <- predict(ss.model,newdata=grid,se=T)$fit
grid$se <- predict(ss.model,newdata=grid,se=T)$se.fit


ggplot(grid,aes(x=time,color=type,group=type)) +
  xlab("Time (ms)") + ylab("Normalized F3 value") +
  geom_line(aes(y=fit,color=type),size=1.5) +
  geom_ribbon(aes(ymin=fit-(1.96*se),ymax=fit+(1.96*se),fill=type),alpha=0.5, color="NA")

```

# Stats - Linear mixed effect models
```{r}
install.packages("lme4")
install.packages("lmerTest")
install.packages("ggplot2")
install.packages("relaimpo")


library(lme4)
library(lmerTest)
library(ggplot2)
library(relaimpo)

# check the first few rows to make sure everything looks fine
head(dat)
tail(dat)

# standarize the measurements, normalized by speaker

  
slist <- c('F03','F04','F05','M06','M07','M08')
for (s in list(dat$speaker)){
  subdata <- dat[dat$speaker==s,]
  
  dat$deltaz[dat$speaker==s] <- scale(subdata$delta,center=T,scale=T)
  dat$f1z[dat$speaker==s] <- scale(subdata$f1,center=T,scale=T)
  dat$f2z[dat$speaker==s] <- scale(subdata$f2,center=T,scale=T)
  dat$f3z[dat$speaker==s] <- scale(subdata$f3,center=T,scale=T)
}  


# check no. of rows
nrow(dat)

# attach data
attach(dat)

# scatter plot
plot(samp_norm,deltaz)

# reduce the data to contain only data points that are equal to or less than 3 stdev
dat <- dat[abs(dat$deltaz)<=3 &
                         abs(dat$f1z)<=3 &
                         abs(dat$f2z)<=3 &
                         abs(dat$f3z)<=3,] 

# check no. of rows
nrow(dat)

# stats models
mod.f1 <- lmer(f1z ~ deltaz*vowel*syllable + (1|speaker) + (1|item), data=dat)
mod.f2 <- lmer(f2z ~ deltaz*vowel*syllable + (1|speaker) + (1|item), data=dat)
mod.f3 <- lmer(f3z ~ deltaz*vowel*syllable + (1|speaker) + (1|item), data=dat)

# Bonferroni correction (standard a level/the number of models)
0.05/3
# 0.0167

# check the results of your models
summary(mod.f1)
summary(mod.f2)
summary(mod.f3)


# visualization

# Testing - reordering within facet - some ideas
d <- data.frame(
  f = rep(c("f1", "f2"), each = 3),
  x = rep(c("a", "b", "c"), 2),
  y = c(1, 2, 3, 3, 2, 1)
)

ggplot(d, aes(x, y)) +
  geom_col() +
  facet_wrap(~ f)

# to change order of facet wrap -try this one - it works
dat$syllable <- factor(dat$syllable, levels = c("onset", "coda"))

ggplot(dat,aes(x=deltaz, y=f1z, col=vowel, group=vowel)) +
  geom_point(alpha=0.2) + geom_smooth(method = "lm") +
  facet_wrap(~syllable) + theme_bw()

ggplot(dat,aes(x=deltaz, y=f2z, col=vowel, group=vowel)) +
  geom_point(alpha=0.2) + geom_smooth(method = "lm") +
  facet_wrap(~syllable) + theme_bw()

ggplot(dat,aes(x=deltaz, y=f3z, col=vowel, group=vowel)) +
  geom_point(alpha=0.2) + geom_smooth(method = "lm") +
  facet_wrap(~syllable) + theme_bw()

```

# Stats - Relative importance analysis
```{r}
lm1 <- lm(deltaz ~ f1z + f2z + f3z , data=dat)
rel.weights <- calc.relimp(lm1, type = c("lmg"), rela = T)
rel.weights

```


